---
title: "example 3"
author: "Darshini Parekh"
date: "2025-12-08"
output: pdf_document
---
library(tidyverse)
library(ggplot2)
```
```{r}
getwd()
setwd("")
data3 <- read.csv("bread dough.csv")
data3
# Convert to factors
data3$Temp <- factor(data3$Temp)
data3$Humidity   <- factor(data3$Humidity)
str(data3)
```
```{r}
summary_temp <- data3 %>%
  group_by(Temp) %>%
  summarise(
    N = n(),
    Mean = mean(Time),
    SD = sd(Time)
  )

summary_hr <- data3 %>%
  group_by(Humidity) %>%
  summarise(
    N = n(),
    Mean = mean(Time),
    SD = sd(Time)
  )

summary_interaction <- data3 %>%
  group_by(Temp, Humidity) %>%
  summarise(
    N = n(),
    Mean = mean(Time),
    SD = sd(Time)
  )
```
```{r}
anova_model <- aov(Time ~ Temp * Humidity, data = data3)
summary(anova_model)
shapiro.test(residuals(anova_model))
qqnorm(residuals(anova_model)); qqline(residuals(anova_model))
leveneTest(Time ~ Temp * Humidity, data = data3)
data3$group <- interaction(data3$Temp, data3$Humidity)

tukey <- TukeyHSD(aov(Time ~ group, data = data3))
print(tukey)

TukeyHSD(anova_model, "Temp")
TukeyHSD(anova_model, "Humidity")
TukeyHSD(anova_model, "Temp:Humidity")
str(data3)
```
Plots
ggplot(data3, aes(x = Temp, y = Time, color = Humidity, group = Humidity)) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  scale_y_reverse() +   # flip Y axis so lower time appears higher = visually better
  theme_minimal() +
  labs(
    title = "Interaction Plot: Effect of Temperature and Humidity on Rise Time",
    x = "Temperature (°C)",
    y = "Mean Rise Time — Time (min)",
    color = "Humidity (%)"
  )
summary_interaction <- data3 %>%
  group_by(Temp, Humidity) %>%
  summarise(
    N    = n(),
    Mean = mean(Time),
    SD   = sd(Time),
    SE   = SD / sqrt(N),
    .groups = "drop"
  )

ggplot(summary_interaction, 
       aes(x = factor(Temp), y = Mean, fill = factor(Humidity))) +
  geom_col(position = position_dodge(0.9)) +
  geom_errorbar(
    aes(ymin = Mean - SD, ymax = Mean + SD),
    position = position_dodge(0.9),
    width = 0.2
  ) +
  scale_fill_brewer(palette = "Blues", direction = -1) +
  coord_cartesian(ylim = c(150, 205)) +  # zoom in to show differences clearly
  theme_minimal() +
  labs(
    title = "Mean Rise Time by Temperature and Humidity",
    x     = "Temperature (°C)",
    y     = "Mean Rise Time — Time (min)",
    fill  = "Humidity (%)",
    caption = "Lower bars = faster rise = better fermentation condition"
  )

# Temperature main effect
ggplot(summary_temp, aes(x = Temp, y = Mean)) +
  geom_col(fill = "steelblue", width = 0.5) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = 0.2, linewidth = 0.8) +
  coord_cartesian(ylim = c(170, 200)) +
  theme_minimal() +
  labs(
    title   = "Effect of Temperature on Mean Rise Time",
    x       = "Temperature (°C)",
    y       = "Mean Rise Time — Time (min)",
    caption = "Lower bar = faster rise"
  )


# Humidity main effect
ggplot(summary_hr, aes(x = Humidity, y = Mean)) +
  geom_col(fill = "seagreen", width = 0.5) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = 0.2, linewidth = 0.8) +
  coord_cartesian(ylim = c(170, 200)) +
  theme_minimal() +
  labs(
    title   = "Effect of Humidity on Mean Rise Time",
    x       = "Relative Humidity (%)",
    y       = "Mean Rise Time — Time (min)",
    caption = "Lower bar = faster rise"
  )
# Temperature main effect
ggplot(summary_hr, aes(x = Temp, y = Mean)) +
  geom_col(fill = "steelblue", width = 0.5) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = 0.2, linewidth = 0.8) +
  coord_cartesian(ylim = c(170, 200)) +
  theme_minimal() +
  labs(
    title   = "Effect of Temperature on Mean Rise Time",
    x       = "Relative Temperature (%)",
    y       = "Mean Rise Time — Time (min)",
    caption = "Lower bar = faster rise"
  )
ggplot(data3, aes(x = interaction(Temp, Humidity, sep = "°C × "),
                  y = Time,
                  fill = factor(Humidity))) +
  geom_boxplot(alpha = 0.7) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Distribution of Rise Time by Treatment Combination",
    x     = "Temperature × Humidity",
    y     = "Rise Time — Time (min)",
    fill  = "Humidity (%)"
  )
```










